\bf The forestecology R package for fitting and assessing neighborhood models of the effect of interspecific competition on the growth of trees

### 2021-07-26 paper/paper.tex
compare multiple competitive distances

Above we set the cut off distance for two stems to compete at 7.5 m. Our
package can be used to find which distance is best supported by the
data. Here we give an example.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mult_dist_comp <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
  \DataTypeTok{dist =} \KeywordTok{c}\NormalTok{(}\FloatTok{2.5}\NormalTok{, }\DecValTok{5}\NormalTok{, }\FloatTok{7.5}\NormalTok{, }\DecValTok{10}\NormalTok{, }\FloatTok{12.5}\NormalTok{),}
  \DataTypeTok{rmse =} \DecValTok{0}
\NormalTok{)}

\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(mult_dist_comp}\OperatorTok{$}\NormalTok{dist))}
\NormalTok{\{}
\NormalTok{  comp_dist <-}\StringTok{ }\NormalTok{mult_dist_comp}\OperatorTok{$}\NormalTok{dist[i]}
\NormalTok{  growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}

\NormalTok{  focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{)}
\NormalTok{  focal_vs_comp_scbi }\OperatorTok{%>%}\StringTok{ }
\StringTok{    }\KeywordTok{select}\NormalTok{(focal_ID, focal_sp, geometry, growth, comp)}

\NormalTok{  focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{focal_vs_comp_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}

@@

>>>>

For all the above analyses we took the cut off distance
(\texttt{comp\_dist}) for two stems to compete as 7.5 m. This distance
has been estimated in previous neighborhood competition studies in
forests \citep[\citet{uriarte_spatially_2004},
\citet{canham_neighborhood_2006}]{canham_neighborhood_2004}. We used 7.5
m in \citet{allen_permutation_2020} as an average of the values
estimated in other studies. But our package can be used to find which
distance is best supported by the data. Here we provide an example using
another section of the SCBI plot to provide an additional example of the
CV block layout. To speed computation we do not consider species
differences in competitive effects and treat all species as the same.

\begin{Shaded}
\begin{Highlighting}[]

\NormalTok{census_}\DecValTok{2013}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem2.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{, gy }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{)}

\NormalTok{census_}\DecValTok{2018}\NormalTok{_scbi <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{"scbi.stem3.csv"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(stemID, sp, }\DataTypeTok{date =}\NormalTok{ ExactDate, gx, gy, dbh, codes, status) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{date =} \KeywordTok{mdy}\NormalTok{(date),}
    \DataTypeTok{dbh =} \KeywordTok{as.numeric}\NormalTok{(dbh)}\OperatorTok{/}\DecValTok{10}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(gx }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{, gy }\OperatorTok{<}\StringTok{ }\DecValTok{400}\NormalTok{)}

\NormalTok{growth_scbi <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
    \DataTypeTok{id =} \StringTok{"stemID"}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\CommentTok{# make all species the same, needs to be factor with at least two levels}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =} \KeywordTok{factor}\NormalTok{(}\StringTok{'A'}\NormalTok{, }\DataTypeTok{levels =} \KeywordTok{c}\NormalTok{(}\StringTok{'A'}\NormalTok{, }\StringTok{'B'}\NormalTok{)) )}

\NormalTok{study_region_scbi <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
  \DataTypeTok{x =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{),}
  \DataTypeTok{y =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{0}\NormalTok{)}
\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{()}

\NormalTok{n_fold <-}\StringTok{ }\DecValTok{4}
\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{))}
\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{200}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{))}
\NormalTok{fold3 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{0}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{))}
\NormalTok{fold4 <-}\StringTok{ }\KeywordTok{cbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{200}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{200}\NormalTok{, }\DecValTok{400}\NormalTok{, }\DecValTok{400}\NormalTok{))}

\NormalTok{growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ spatial_block_scbi}\OperatorTok{$}\NormalTok{foldID }\OperatorTok{%>%}\StringTok{ }\KeywordTok{factor}\NormalTok{())}

\NormalTok{mult_dist_comp <-}\StringTok{ }\KeywordTok{tibble}\NormalTok{(}
  \DataTypeTok{dist =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{ ,}\FloatTok{6.25}\NormalTok{,  }\FloatTok{7.5}\NormalTok{, }\FloatTok{8.75}\NormalTok{,  }\DecValTok{10}\NormalTok{),}
  \DataTypeTok{rmse =} \DecValTok{0}
\NormalTok{)}

\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\KeywordTok{length}\NormalTok{(mult_dist_comp}\OperatorTok{$}\NormalTok{dist))}
\NormalTok{\{}
\NormalTok{  comp_dist <-}\StringTok{ }\NormalTok{mult_dist_comp}\OperatorTok{$}\NormalTok{dist[i]}

\NormalTok{  growth_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{size =}\NormalTok{ comp_dist, }\DataTypeTok{region =}\NormalTok{ study_region_scbi)}

\NormalTok{  focal_vs_comp_scbi <-}\StringTok{ }\NormalTok{growth_scbi }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi, }\DataTypeTok{id =} \StringTok{"stemID"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{run_cv}\NormalTok{(}\DataTypeTok{comp_dist =}\NormalTok{ comp_dist, }\DataTypeTok{blocks =}\NormalTok{ blocks_scbi)}

@@

Here we see that a cut off distance of approximately 6 m minimizes RMSE.
